\title{Paper Draft}
\author{Henry et al}
\date{} % keep if you want to suppress date

\maketitle

\section*{LaTeX Help and Doc Standards}
Example sentence that mentions \Figref{fig:arenafig}. Example sentence about some data (\figref{fig:arenafig}). Note how the refs are handled differently.

\section{Introduction}

\hl{From F31 resubmission (obviously this needs some editing/reworking in light of new stuff)}

The inevitability of aging makes it unique among biological challenges. Nearly all organisms experience physical deterioration as they age, and aging itself is a primary risk factor for many diseases, including cancer, diabetes, cardiovascular disease, and neurodegeneration.1 Though aging was long assumed to be a natural byproduct of lifelong wear-and-tear, we now know that lifespan can be positively influenced by the manipulation of an organism’s environment. The most well-known example of this is dietary restriction (DR), which impacts health and longevity through highly conserved cell- and tissue-autonomous mechanisms. A cell-autonomous explanation is intuitive, but we know that this picture of a bottom-up response to the environment is incomplete. It is now well established that the nervous system mediates a significant component of the lifespan phenotypes of several environmental manipulations.2,3 For example, multiple model systems have been used to demonstrate that merely perceiving food odors while on a restricted diet mitigates the benefits of DR, while the inability to detect food odors extends lifespan on an ad-libitum diet.4–6 In addition to smell and taste, our lab has shown that visual perception is necessary and sufficient to modulate lifespan in flies housed with dead conspecifics.7 These and other examples demonstrate that the brain plays a significant role in regulating aging.

How the brain influences aging in flies and other taxa is poorly understood. Our lab aims to uncover the mechanisms by which environmental information is transduced into different aging outcomes. Does the brain “prime” a peripheral response to anticipated inputs, initiating or amplifying metabolic processes in target tissues? Does the experience per se of a cue induce or relieve a neural state in the animal, such as a state related to stress, fear, or frustration, that modulates lifespan? Are these changes physiological consequences of behavioral decisions? Evidence exists in various contexts that is consistent with each of these possibilities.7–10 Whether neurally-mediated lifespan phenotypes arise through distinct or common mechanisms is also an open question. The latter possibility is especially intriguing—just as slowing aging offers an upstream approach to combating many physiological problems simultaneously, neural regulation of aging represents a potential upstream point of intervention in multiple aging processes. 

Identifying neural structures that mediate multiple aging phenotypes would be a promising step toward this ideal. Our lab has determined that the ellipsoid body (EB) is a candidate for such a structure in the Drosophila brain. The EB is a component of the central complex (CX), a region functionally and structurally analogous to the mammalian basal ganglia (BG).11 The EB has roles in sensory integration,12 memory formation,13 sleep regulation,14 motor responses,12,15 and other fundamental processes that allow a fly to perceive and respond to its environment. While investigating the mechanisms underlying the death perception lifespan phenotype described above, we discovered that silencing the EB or a few of its constituent ring neuron populations—namely, R2 and R4 neurons—protects flies from the accelerated aging effects of death perception.16 Furthermore, we have discovered that direct manipulation of these neurons is sufficient to extend lifespan: specifically, activation of R2 ring neurons (Fig. 1) or silencing of a subset of R4 neurons, R4d (Fig. 2). Intriguingly, these neurons have also been implicated in normal modulatory roles involving health, behavior, and aging, demonstrating their importance for physiological regulation and their ability to influence lifespan in natural contexts. For example, in addition to protecting flies from the negative effects of death perception, silencing R4d rescues the lifespan of flies separately presented17,18 with sugar and protein dietary components (termed a “choice diet”) (Fig. 3), showing that the accelerated aging phenotype requires these neurons. R4d activity also promotes survival under nutrient stress19 and drives changes in behavior and physiology associated with different levels of dietary amino acids (not shown). Considering these observations and the EB’s known role as an integrative center, we posit that these neurons play important roles in influencing health and longevity outcomes in response to natural environmental and internal cues. This offers a proof of principle that lifespan-modulating neurons can be beneficially manipulated, and since this structure has a mammalian analogue, understanding more about these neurons can inform us about similar targets in mammals. 

Our investigations of the death perception and choice diet mechanisms hint that different stimuli recruit R2 and R4d through common pathways, including serotonin signaling pathways, suggesting that these neurons have evolved to play a key role in processing environmental cues. Knocking down serotonin receptor 5-HT2A in R2 and R4d abolishes the lifespan phenotype of death perception,16 and knocking down 5-HT2A in R4d also rescues lifespan in flies fed a choice diet (Fig. 4) (R2 not yet studied). These findings indicate that serotonergic inputs may serve as an evolutionarily conserved system that bridges environmental information and aging outputs. However, we still know little about how these neurons behave under normal conditions to influence health and lifespan, and what the downstream targets of their activity might be. 

Our goal in this proposal is to investigate mechanisms by which these neurons modulate health and aging in Drosophila, which will elucidate principles underlying the neural control of aging. This work will provide a starting point for tracing the effects of these neural signals to their downstream targets, which will in turn reveal whether the processes involved are conserved across taxa. Because the EB and CX have mammalian analogues, we expect that understanding these neurons and their influence on physiology will give us a framework for identifying similar patterns and mechanisms in mammalian brains. Most importantly, it will suggest precise upstream targets for intervention—an important step toward slowing aging in humans. 


