%% Just collecting methods from other people here for reference as I go through my sections; need to modify them 

\subsection{Fly stocks and husbandry}

%% Kristy Science paper
Fly stocks were maintained on a standard cornmeal-based larval growth medium and in a controlled environment (21C, 60\% humidity) with a 12:12hr light:dark cycle. We controlled the developmental larval density by manually aliquoting 32 ul of collected eggs into individual bottles containing 25-50mL of food at 25C. Following eclosion, mixed sex flies were kept on SY10\% medium (10\% w/v sucrose and 10\% w/v yeast) for 2-4 days until they were sorted by sex and transferred onto chemically defined food or kept on SY10\% food, as needed, for experiments. Experimental flies were flipped to fresh food e/o day until completion of the experiment. Unless otherwise noted, we used mated female flies that were between 7-12 days old for all experiments. The following stocks were used for experiments: Canton-S and w 1118 were obtained from the Bloomington Drosophila Stock Center. UAS-CsChrimson (BDSC #55135 and #55136), UAS-H3.3A-GFP (BDSC #68241), UAS-HIRA RNAi (BDSC #35346), GMR50H05-GAL4 (BDSC #38764), Trh-GAL4 (BDSC #38388), UAS-BCAT RNAi (VDRC KK110229), UAS-BCAT RNAi (VDRC GD25204), and Rpd3def24 (BDSC 32239) were purchased from BDSC or the Vienna Drosophila Resource Center, as indicated. PoxN mutants were provided by J. Alcedo (Wayne State University). Nsyb-GAL4 was provided by L. Buttitta (University of Michigan, MI) and Nsyb-GeneSwitch-GAL4 was obtained from A. Sehgal (Perelman School of Medicine, PA). UAS-GtACR was provided by M. Dus (University of Michigan, MI). All transgenic lines used in this study, with the exception of UAS-H3.3A-GFP, UAS-HIRA RNAi, Nsyb-GAL4, Nsyb-GeneSwitch-GAL4, and Rpd3def24 were back-crossed 10 generations to w-;Canton-S prior to experiments

Lifespan measurements74
Lifespans were measured using established protocols (69). We established 6-10 replicate vials75
for each treatment, with 20 flies per vial. Flies were transferred to fresh media every 2-3 days, at76
which time dead flies were removed and recorded using the DLife system (69). Flies were kept77
in constant temperature (25C) and humidity (60%) conditions with a 12:12hr light:dark cycle.78
Normally, we conducted at least two experimental replicates of each lifespan experiment (see79
Table S3).80
81
Optogenetic Assays82
Flies expressing UAS-CsChrimson in desired neuronal populations were reared as described83
above. Upon eclosion, females were sorted to SY10% medium containing 800 μm all-trans-84
retinal (from a stock solution of 100mM ATR in 100% ethanol) and kept in the dark for 2 days.85
Flies were then flipped to appropriate holidic diet food containing 400 μm ATR (and flipped to86
fresh ATR-containing food e/o day for the duration of the experiment) and moved to a custom87
rig containing 627nm (red) LEDs or 530nm (green) LEDs (Luxeon) or kept in the dark in the88
same incubator (Fig. S3). The custom rig is fully enclosed to prevent leakage of light and houses89
48 individual vials surrounded by mirrors. Custom hardware and firmware were designed to90
allow the experimenter to control the LED intensity and a range of other light stimulus91
parameters. Lifespan and Con-Ex experiments used a stimulus frequency of 40 Hz and a pulse92
width of 800 ms. Flies were exposed to this protocol every day for 12 hours per day during the93
light period, followed by 12 hours of darkness.94
95
Consumption-excretion feeding assays96
Con-Ex experiments were carried out as previously described (25, 26). Experimental female flies97
were sorted to appropriate BCAA diets (10 flies per vial, 8-10 replicates for each treatment)98
when they were >4 days old. Food was changed every 2-3 days until the appropriate99
experimental time-point was reached (5 days for standard experiments, unless otherwise100
indicated). After the dietary pre-treatment period, blue test food was prepared in removable caps101
by adding 1% (w/v) FD&C Blue No. 1 to the appropriate diets. Flies were moved to fresh vials102
with the removable blue-food caps on the top of the vials and were allowed to feed and excrete103
for 24 hours. Caps and flies were removed after the 24-hour test period and flies were counted.104
Excreted dye was collected by vortexing each vial with 3mL of water. Concentration of the dye105
was determined by absorbance at 630nm and compared to a standard curve of known106
concentrations. For short-term Con-Ex experiments lasting <3 hours, both excretions and flies107
were collected and measured. To measure internal blue concentration, groups of flies were108
homogenized in 1mL of phosphate-buffered saline containing 0.1% Trixon X-100 (IBI109
Scientific) for 30s at 30Hz using a QIAGEN TissueLyser. Concentration of internal extracts110
were determined by absorbance at 630nm and both internal and excreted concentrations were111
summed to determine total consumption.